Estimate Response to Selection

### 2021-06-24 R/03-response.Rmd
convert response to selection script to rmd

This script intakes the flux data from script 01. It then calculates deviance
as the difference between the positive line and the neutral line in each
passage. The errors can either be calculated as the sum of the errors for
each line or the standard error of difference, that is, the square root of
the sum of the squared errors, which is slightly smaller than the sum. In
this case, I have chosen the SED.
```{r initialize_packages, message=FALSE}
library(tidyverse)
library(morris)
library(broom)
library(knitr)
```

@@

```{r import_data, message=FALSE}
fluxes <- read_tsv(paste0(der_dir, 'fluxes.tsv'))
```

@@

```{r calc_plot_ratios}

```{r plot_raw}
ratio <- calc_plot_ratios(fluxes$passage, fluxes$estimate)

plot_flux(fluxes, passage, estimate, ratio)
#ggsave('fluxes_raw.pdf', width = 5, height = 4)
```

@@

```{r plot_log10}
ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$estimate_log10)

plot_flux(fluxes, passage, estimate_log10, ratio_log10, log10 = TRUE)

@@

```{r response_model}
# Fit full interaction model
inter_model <- lm(estimate_log10 ~ 0 + passage * treat, data = fluxes)

# Fit additive model
addit_model <- lm(estimate_log10 ~ passage + treat, data = fluxes)

# Back calculate the slope of the positive treatment
pos_slope <- 10^inter_model$coef["passage:treatp"]
```

@@

```{r compare_models}
# Compare interaction to no interaction
compare_mods <- anova(inter_model, addit_model)
kable(tidy(compare_mods))
```

@@

```{r inter_model}
kable(tidy(inter_model))
```

The term `passage` was not significant, which is the y-intercept of the neutral treatment So This treatment did not change.

@@

Previously, I was using the `metafor` package to test the slope of the difference between the positive and neutral line.

>>>>

This script intakes the flux data from script 01. It then fits a linear model to the flux data over time per treatment and tests a difference of slope between the positive and the neutral treatment. Finally, it 
```{r initialize_packages, message=FALSE}
library(tidyverse)
library(morris)
library(broom)
library(knitr)
#library(broman)
```

@@

```{r import_data, message=FALSE}
fluxes <- read_tsv(paste0(der_dir, 'fluxes.tsv'))

fluxes <-
  fluxes %>% 
  mutate(passage = passage - 1)
```

@@

```{r plot_raw}
ratio <- calc_plot_ratios(fluxes$passage, fluxes$ch4)

plot_flux(fluxes, passage, ch4, ratio)
#ggsave('fluxes_raw.pdf', width = 5, height = 4)
```

@@

```{r plot_log10}
ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$log_ch4)

plot_flux(fluxes, passage, log_ch4, ratio_log10, log10 = TRUE)

@@

```{r response_model}
# Fit full interaction model
inter_model <- lm(log_ch4 ~ passage * treat, data = fluxes)

# No treatment model
no_treat_model <- lm(log_ch4 ~ passage, data = fluxes)

# Back calculate the slope of the positive treatment
pos_slope <- 10^inter_model$coef["passage:treatp"]

@@

```{r compare_models}
# Compare interaction to no interaction
compare_mods <- anova(inter_model, no_treat_model)
kable(tidy(compare_mods))
```

@@

```{r inter_model}
kable(tidy(inter_model))

The intercept for the Neutral treatment was significantly different from zero at `r round(tidy(inter_model)[[1, 2]], 2)` (SE = `r round(tidy(inter_model)[[1, 3]], 2)`, t = `r round(tidy(inter_model)[[1, 4]], 2)`, p = `r round(tidy(inter_model)[[1, 5]], 2)`). The Positive treatment started at a significantly lower methane oxidation rate than the Neutral treatment with a y-intercept of `r round(tidy(inter_model)[[3, 2]], 2)` relative to Neutral (SE = `r round(tidy(inter_model)[[3, 3]], 2)`, t = `r round(tidy(inter_model)[[3, 4]], 2)`, p = `r round(tidy(inter_model)[[3, 5]], 2)`). There was no change in methane oxidation rate of the Neutral treatment over the five passages (slope = `r round(tidy(inter_model)[[2, 2]], 2)`, SE = `r round(tidy(inter_model)[[2, 3]], 2)`, t = `r round(tidy(inter_model)[[2, 4]], 2)`, p = `r round(tidy(inter_model)[[2, 5]], 2)`). By contrast, there was a significant increase in the Positive treatment of `r round(tidy(inter_model)[[4, 2]], 2)` per passage which is approximately a `r round((pos_slope - 1) * 100, 0)`% increase in methane oxidation rate per passage (SE = `r round(tidy(inter_model)[[4, 3]], 2)`, t = `r round(tidy(inter_model)[[4, 4]], 2)`, p = `r round(tidy(inter_model)[[4, 5]], 2)`).

@@

Previously, I was using the `metafor` package to test the slope of the difference between the positive and neutral line. For this model, I calculate the difference between the Positive and Neutral treatment and fit that difference as the response. I also calculate the standard error of the difference and included that variance in the model. I then fit the model `difference ~ passage` to test the slope of the difference over time.
